Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-09 paper.tex
oops

\graphicspath{{"figures/"}}
\begin{document}
%Title of paper
\title{Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies}

@@

To employ the Albrecht formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a) \otimes \ket{b}$ for electronic states $\{|a)\}$ and vibrational states $\{\ket{b}\}$. \cite{BornOppenheimer, Albrecht1960}
For simplicity, we choose $\ket{g}, \ket{v}$ to be vibrational states on the ground electronic state $|g)$ and $\ket{m} \rightarrow \ket{m,j}$, $\ket{n} \rightarrow \ket{n,k}$ are virtual states.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,j,k} \frac{
		R_{i}^{gv, nk} 
		R_{j}^{nk,mj} 
		R_{k}^{mj,gg} 
		R_{l}^{gv,gg} 
	}{\Delta_{g0,gg}
		\Delta_{nk, mj}
		\Delta_{mj, gg}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.\cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.\cite{Albrecht1960}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{M_{l}^{gg}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z}
\end{equation} yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Ziegler1988} so that
\begin{equation}
		\gamma_{ijkl} \sim \frac{(A_{ijk} + B_{ijk} + C_{ijk})\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term depends upon Franck-Condon ($\order{Q^0}$) transitions, the $B$ term depends upon Herzberg-Teller ($\order{Q}$) transitions, and the $C$ term depends on $\order{Q^2}$ transitions. 
The C term is suppressed in the following discussion as it depends on one and two photon forbidden transitions which have yet to be observed in spontaneous hyper-Raman scattering. \cite{Ziegler1988}
In the discussion that follows, we take $\hbar = e = 1$.

@@

When $\omega_2$ and $\omega_3$ are largely detuned from a local resonance, the resonance denominators lose their dependence on vibrational quantum number.
In the case of the $A$ term, the sum can now be contracted over ${j, k}$. 
Assuming a normal mode basis, $A_{ijk} \sim \sum_{j,k}\langle v | k \rangle \langle k | j \rangle \langle j | g \rangle = \delta_{vg} = 0$, or, A$_{ijk}$ vanishes for singly resonant HDFG. \cite{Neddersen1989}
The B terms do not suffer from this problem.
The Herzberg-Teller matrix elements, where $\ket{z} \neq \{\ket{g}, \ket{v}\}$ is a normal mode, allows for non-zero B terms.
As a result, singly resonant HDFG has significant dependence upon those modes which participate in Herzberg-Teller coupling. 
This likely explains the relative weakness of HDFG response from hexane $\nu$(CH) modes. \cite{RN351}
The singly resonant HDFG susceptibility can then be approximated as
\begin{equation}
	\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, large $B$ values are a necessity for singly resonant HDFG response.
\subsection{Doubly Resonant}
While HDFG can provide useful information about the properties of molecular vibrations, it is apparent that HDFG can inform on molecular electronic structure by making the hyper-Raman transition resonant with a vibronic state, i.e. doubly resonant HDFG (DR-HDFG).
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling, analogous to fully resonant DFG. \cite{Dick83_1, Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.
Two beam, parametric and non-parametric DR-HDFG processes are diagrammed in \autoref{fig:doubsive}.

>>>>

\graphicspath{{"figures/"}}
\begin{document}
%Title of paper
\title{Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopy}

@@

To employ the Albrecht formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a) \otimes \ket{b}$ for electronic states $\{|a)\}$ and vibrational states $\{\ket{b}\}$. \cite{BornOppenheimer, Albrecht1960}
For simplicity, we choose $\ket{g}, \ket{v}$ to be vibrational states on the ground electronic state $|g)$ and $\ket{m} \rightarrow \ket{m,j}$, $\ket{n} \rightarrow \ket{n,k}$ are virtual states.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,j,k} \frac{
		R_{i}^{gv, nk} 
		R_{j}^{nk,mj} 
		R_{k}^{mj,gg} 
		R_{l}^{gv,gg} 
	}{\Delta_{g0,gg}
		\Delta_{nk, mj}
		\Delta_{mj, gg}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.\cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.\cite{Albrecht1960}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{M_{l}^{gg}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z
\end{equation} 
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Ziegler1988} so that
\begin{equation}
		\gamma_{ijkl} \sim \frac{(A_{ijk} + B_{ijk} + C_{ijk})\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term depends upon Franck-Condon ($\order{Q^0}$) transitions, the $B$ term depends upon Herzberg-Teller ($\order{Q}$) transitions, and the $C$ term depends on $\order{Q^2}$ transitions. 
The C term is suppressed in the following discussion as it depends on one and two photon forbidden transitions which have yet to be observed in spontaneous hyper-Raman scattering. \cite{Ziegler1988}
In the discussion that follows, we take $\hbar = e = 1$.

@@

When $\omega_2$ and $\omega_3$ are largely detuned from a local resonance, the resonance denominators lose their dependence on vibrational quantum number.
In the case of the $A$ term, the sum can now be contracted over ${j, k}$. 
Assuming a normal mode basis, $A_{qrs} \sim \sum_{j,k}\langle v | k \rangle \langle k | j \rangle \langle j | g \rangle = \delta_{vg} = 0$, i.e., A$_{qrs}$ vanishes for singly resonant HDFG. \cite{Neddersen1989}
The Herzberg-Teller matrix elements, where $\ket{z} \neq \{\ket{g}, \ket{v}\}$ is a normal mode, do not vanish under these conditions.
As a result, the singly resonant HDFG susceptibility can be approximated as
\begin{equation}
	\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, vibronic activity is a necessity for singly resonant HDFG response.
\subsection{Doubly Resonant}
While HDFG can provide useful information about the properties of molecular vibrations, it is apparent that HDFG can inform on molecular electronic structure by making the hyper-Raman transition resonant with a vibronic state, i.e. doubly resonant HDFG (DR-HDFG).
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling, analogous to fully resonant DFG. \cite{Dick83_1, Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.
Two beam, parametric and non-parametric DR-HDFG processes are diagrammed in \autoref{fig:doubsive}.
